---
title: "cellmig: quantifying cell migration with hierarchical Bayesian models"
author: "Simo Kitanovski (simo.kitanovski@uni-due.de)"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{User Manual: cellmig}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE, warning = FALSE}
knitr::opts_chunk$set(comment = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```


# Background

```{r}
library(cellmig)
library(ggplot2)
library(ggforce)
library(patchwork)
ggplot2::theme_set(new = theme_bw(base_size = 10))
```


# Data (raw)
In this vignette we will use simulated data: 
  
  * 6 compounds (chemical): c-1, c-2, ..., c-6
  * 5 doses: 1, 2, 3, 4 and 5
  * 3 sample: s-1, s-2 and s-3
  * 2 batches: b-1 and b-2

In each sample from a combination of compound x dose x batch, we have 
20 observations (cell velocities $v$)

```{r}
data("d", package = "cellmig")
knitr::kable(d[1:10,], row.names = F, digits = 2)
```

```{r, fig.width=7, fig.height=5}
ggplot(data = d)+
  facet_wrap(facets = ~paste0("compound=", compound), scales = "free_y", ncol = 2)+
  geom_sina(aes(x = as.factor(dose), col = plate, y = v, group = sample), size = 0.5)+
  theme_bw()+
  theme(legend.position = "top",
        strip.text.x = element_text(margin = margin(0.03,0,0.03,0, "cm")))+
  xlab(label = "dose")+
  ylab(label = "migration speed")+
  scale_color_manual(values = c("steelblue", "orange", "forestgreen", "purple"))+
  guides(color = guide_legend(override.aes = list(size = 3)))
```


# Migration speed (scaled)

```{r, fig.width=7, fig.height=5}
ggplot(data = d)+
  facet_wrap(facets = ~paste0("compound=", compound), scales = "free_y", ncol = 2)+
  geom_sina(aes(x = as.factor(dose), col = plate, y = v/max(v), group = sample), size = 0.5)+
  theme_bw()+
  theme(legend.position = "top",
        strip.text.x = element_text(margin = margin(0.03,0,0.03,0, "cm")))+
  xlab(label = "dose")+
  ylab(label = "migration speed [scaled by max speed]")+
  scale_color_manual(values = c("steelblue", "orange", "forestgreen", "purple"))+
  guides(color = guide_legend(override.aes = list(size = 3)))
```


# Model fitting

```{r, fig.width=7, fig.height=3.5}
o <- cellmig(x = d,
             control = list(mcmc_warmup = 300,
                            mcmc_steps = 1300,
                            mcmc_chains = 4,
                            mcmc_cores = 4,
                            mcmc_algorithm = "NUTS",
                            adapt_delta = 0.8,
                            max_treedepth = 10))
```


# What are the overall effects of compounds and doses on cell migration?

```{r, fig.width=6, fig.height=3}
ggplot(data = o$s$mu_group)+
  geom_line(aes(x = dose, y = mean, col = compound, group = compound))+
  geom_point(aes(x = dose, y = mean, col = compound))+
  geom_errorbar(aes(x = dose, y = mean, ymin = X2.5., ymax = X97.5., 
                    col = compound), width = 0.04)+
  ylab(label = "effect")
```



# Compare the dose-response `profiles` for different compounds

```{r, fig.width=9, fig.height=5}
get_profiles(x = o)+
  patchwork::plot_layout(widths = c(1, 1, 5))
```


# Compare the effects between treatment group

## Pairwise dot-plot comparison $\rightarrow$ x minus y axis
  * x/y-axis treatment groups (combinations of compounds and doses)
  * $\delta$: Difference between treatment groups at y-x axis.
  * $\pi$: probability of seing either completely positive or negative $\delta$

```{r, fig.width=8, fig.height=6}
u <- get_pairs(x = o)
u$plot
```



## Violin plot based targeted comparison
  * x/y-axis treatment groups (combinations of compounds and doses)
  * $\delta$: Difference between treatment groups at y-x axis.
  * $\pi$: probability of seing either completely positive or negative $\delta$

```{r, fig.width=7, fig.height=2.5}
u <- get_violins(x = o, 
                 from_groups = c("c1|d1", "c1|d2", "c1|d3", "c1|d4", "c1|d5"), 
                 to_group = "c2|d1")
u$plot
```



# Posterior predictive checks (PPCs)

```{r, fig.width=5, fig.height=5}
g <- get_ppc(x = o)
g
```


# `migcell` can be used to simulate data

```{r}
sim_data <- sim(N_well_cells = 20, 
                N_plate = 3,
                N_group = 6,
                N_well_reps = 3,
                shape = 4,
                sigma_bplate = 0.3,
                sigma_wplate = 0.1,
                alpha_plate = c(1, 1.2, 1),
                mu_group = c(-1, -0.5, -0, 0.5, 1, 2))
```

```{r, fig.width=6, fig.height=3}
ggplot(data = sim_data)+
  geom_sina(aes(x = paste0("group=", group), col = plate, y = v, 
                group = sample), size = 0.5)+
  theme_bw()+
  theme(legend.position = "top",
        strip.text.x = element_text(margin = margin(0.03,0,0.03,0, "cm")))+
  xlab(label = "dose")+
  ylab(label = "migration speed")+
  scale_color_manual(values = c("steelblue", "orange", "forestgreen", "purple"))+
  guides(color = guide_legend(override.aes = list(size = 3)))
```



# Session Info

```{r}
sessionInfo()
```

